Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula

### 2024-02-13 output/paper/manuscript.qmd
Added references biological mechanisms EDCs to cortisol as MV v1.

Our results are consistent with prior epidemiological research that associated exposure to certain {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. There are some differences, though. First, these studies only focus on phthalates, either as individual metabolites or as mixture. Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively. Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023]. Contrary to these studies, we did find effect modification by sex. We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.

>>>>

Our results are consistent with prior epidemiological research that associated exposure to certain {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. There are some differences, though. First, these studies only focus on phthalates, either as individual metabolites or as mixture. Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively. Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023]. Contrary to these studies, we did find effect modification by sex. We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood. Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011]. There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
